Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex

Given a classifier that is able to discriminate on a dataset and identify a certain cognitive state, we may ask ourselves which parts of the brain contain the information leading to this identification. \emph{Pattern localization} addresses this problem by selecting the voxels that contribute the strongest and most reliably to the classification. Surprisingly, when taking the intersection of selected voxels across multiple subjects (for example, when performing cross-validation), this intersection is only a fairly small subset of voxels~\cite{pereira2009machine}.

>>>>

Given a classifier that is able to discriminate on a dataset and identify a certain cognitive state, we may ask ourselves which parts of the brain contain the information leading to this identification. \emph{Pattern localization} addresses this problem by selecting the voxels that contribute the strongest and most reliably to the classification. Surprisingly, when taking the intersection of selected voxels across multiple subjects (for example, when performing cross-validation), this intersection is only a fairly small subset of voxels~\cite{pereira2009machine}. The reason for this is that brains differ between subjects in their size, shape, and location of particular cells. As a result, activation patterns often do not line up well. One of the strategies to deal with this is to use the set of voxels active for each subject as seeds in a clustering algorithm to find similar voxels.\\

In \emph{pattern characterization}, the question of \emph{how} class information is represented takes a central place. While this paper will focus on linear classifiers, some studies in this area resort to a more sophisticated approach and revolve around the question how to correspond relationships in stumulus with those in fMRI data. For example,~\cite{mitchell2008predicting} researches predicting fMRI response for new words based on their similarity to words for which the response is known.
